---
title: "p8105_hw2_baw2160"
author: "Bryce Wong"
date: "September 28, 2018"
output: github_document
---
```{r Set Up, include=FALSE}
library(tidyverse)
library(readxl)
```
##Problem One:

###Read and clean the data: 

Reading in csv file and adjusting the variable names to lowercase snakecase. Retaining line, station name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance. Also converting the entry variable from character (YES vs NO) to a logical variable. 


```{r Reading and Cleaning}
options(tibble.print_min = 3)

NYC_transit_data = read_csv(file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(line, station_name, station_latitude, station_longitude, route1:route11, entry, vending, entrance_type, ada) %>% 
  mutate(entry = recode(entry, "YES" = TRUE, "NO" = FALSE))
```

### Information about the dataset:

This dataset contains information about the NYC transit system. Each observation is every potential station in the city, organized by which line it is on. For each station that is observed, information is included about that station, including location, existence of exits or entrances, and accessibility. Each of these details are variables that the observations can have information about.

To clean the dataset, I've simply converted all variable names to lowercase snakecase and retained a subsection of variables from the original dataset:  

* Where the station is (latitudinally and longitudinally) 
* Whether there is an entrance and what type of entrance it is (stairs, elevator, etc.)
* Whether there is a vendor (a small convenience store-like apparatus)
* Whether the station is ADA accessible, and 
* Which train uses the stations different routes (not all stations have the same number of routes)

The data set has size `r dim(NYC_transit_data)[1]` x `r dim(NYC_transit_data)[2]` and I would consider this...*not* tidy data?

###Answering some questions about the data:

How many distinct stations are there? 

```{r}
NYC_transit_distinct = distinct(NYC_transit_data, line, station_name, .keep_all = TRUE)
```

There are `r dim(NYC_transit_distinct)[1]` distinct stations.

How many stations are ADA compliant? `r length(which(NYC_transit_distinct$ada == "TRUE"))` stations.

What proportion of station entrances / exits without vending allow entrance?

`r length(which((NYC_transit_distinct$vending == "NO") & (NYC_transit_distinct$entry == "TRUE")))` out of `r length(which(NYC_transit_distinct$vending == "NO"))` station entrances/exits without vending allow entrance.

Reformat data so that route number and route name are distinct variables. 
```{r}
NYC_transit_tidy = gather(NYC_transit_distinct, key = route_num, value = route_name, route1:route11)
```

How many distinct stations serve the A train? 

`r length(which(NYC_transit_tidy$route_name == "A"))` distinct stations serve the A train.

Of the stations that serve the A train, how many are ADA compliant?

`r length(which((NYC_transit_tidy$route_name == "A") & (NYC_transit_tidy$ada == "TRUE")))` stations.

## Problem 2:

### Read and clean the Mr. Trash Wheel sheet:

* specify the sheet in the Excel file and to omit columns containing notes (using the range argument and cell_cols() function)
* use reasonable variable names
* omit rows that do not include dumpster-specific data
* rounds the number of sports balls to the nearest integer and converts the result to an integer variable (using as.integer)

```{r}
trashwheel_data = read_excel("./data/HealthyHarborWaterWheelTotals2017-9-26.xlsx", sheet = "Mr. Trash Wheel", range = cell_cols("A:N")) %>%
  janitor::clean_names() %>%
  filter(!is.na(dumpster), month != "Grand Total") %>%
  mutate(sports_balls = as.integer(round(sports_balls)))
```

### Read and clean precipitation data for 2016 and 2017:

For each, omit rows without precipitation data and add a variable year. Next, combine datasets and convert month to a character variable (the variable month.name is built into R and should be useful).

```{r}
precip_2016_data = read_excel("./data/HealthyHarborWaterWheelTotals2017-9-26.xlsx", sheet = "2016 Precipitation", range = "A2:B14") %>%
  janitor::clean_names() %>%
  mutate(year = "2016")
```

```{r}
precip_2017_data = read_excel("./data/HealthyHarborWaterWheelTotals2017-9-26.xlsx", sheet = "2017 Precipitation", range = "A2:B10") %>%
  janitor::clean_names() %>%
  mutate(year = "2017")
```

```{r}
precip_2016_and_2017 = bind_rows(precip_2016_data, precip_2017_data) %>%
  janitor::clean_names() %>%
  mutate(month = month.name[month]) %>% 
  select(year, everything()) 
```

### Talking about the datasets:

* In the Mr. Trash Wheel dataset, there are `r dim(trashwheel_data)[1]` observations. 
* In the combined 2016 and 2017 precipitation datasets, there are `r dim(precip_2016_and_2017)[1]` observations.

Examples of key variables include:

* total preciptation
* month
* year
* dumpster number

For available data, the total precipitation in 2017 was `r sum(precip_2017_data$total)`. The median number of sports balls in a dumpster in 2016 was `r median(subset(trashwheel_data, year == "2016")$sports_balls)`.

## Problem Three:

# install.packages("devtools")
```{r}
devtools::install_github("p8105/p8105.datasets")
library(p8105.datasets)
```

```{r}
brfss_data = brfss_smart2010 %>% 
  filter(Topic == "Overall Health") %>% 
  select(Year, Locationabbr, Locationdesc, Response, Data_value) %>% 
  rename(location = Locationabbr, location_county = Locationdesc, year = Year, response = Response, proportion = Data_value) %>% 
  spread(key = response, value = proportion) %>% 
  janitor::clean_names() %>% 
  mutate(high_health = excellent + very_good)
```

* format the data to use appropriate variable names;
* focus on the “Overall Health” topic
* exclude variables for class, topic, question, sample size, and everything from lower confidence limit to GeoLocation
* structure data so that values for Response (“Excellent” to “Poor”) are column names / variables which indicate the proportion of subjects with each response (which are values of Data_value in the original dataset)
* create a new variable showing the proportion of responses that were “Excellent” or “Very Good”

