---
title: "p8105_hw2_baw2160"
author: "Bryce Wong"
date: "September 28, 2018"
output: github_document
---
```{r Set Up, include=FALSE}
library(tidyverse)
```
##Problem One:

###Read and clean the data: 

Reading in csv file and adjusting the variable names to lowercase snakecase. Retaining line, station name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance. Also converting the entry variable from character (YES vs NO) to a logical variable. 


```{r Reading and Cleaning}
options(tibble.print_min = 3)

NYC_transit_data = read_csv(file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(line, station_name, station_latitude, station_longitude, route1:route11, entry, vending, entrance_type, ada) %>% 
  mutate(entry = recode(entry, "YES" = TRUE, "NO" = FALSE))
```

### Information about the dataset:

This dataset contains information about the NYC transit system. Each observation is every potential station in the city, organized by which line it is on. For each station that is observed, information is included about that station, including location, existence of exits or entrances, and accessibility. Each of these details are variables that the observations can have information about.

To clean the dataset, I've simply converted all variable names to lowercase snakecase and retained a subsection of variables from the original dataset:  

* Where the station is (latitudinally and longitudinally) 
* Whether there is an entrance and what type of entrance it is (stairs, elevator, etc.)
* Whether there is a vendor (a small convenience store-like apparatus)
* Whether the station is ADA accessible, and 
* Which train uses the stations different routes (not all stations have the same number of routes)

The data set has size `r dim(NYC_transit_data)[1]` x `r dim(NYC_transit_data)[2]` and I would consider this...*not* tidy data?

###Answering some questions about the data:

How many distinct stations are there? 

```{r}
NYC_transit_distinct = distinct(NYC_transit_data, line, station_name, .keep_all = TRUE)
```

There are `r dim(NYC_transit_distinct)[1]` distinct stations.

How many stations are ADA compliant? `r length(which(NYC_transit_distinct$ada == "TRUE"))` stations.

What proportion of station entrances / exits without vending allow entrance?

`r length(which((NYC_transit_distinct$vending == "NO") & (NYC_transit_distinct$entry == "TRUE")))` out of `r length(which(NYC_transit_distinct$vending == "NO"))` station entrances/exits without vending allow entrance.

Reformat data so that route number and route name are distinct variables. 
```{r}
NYC_transit_tidy = gather(NYC_transit_distinct, key = route_num, value = route_name, route1:route11)
```

How many distinct stations serve the A train? 

`r length(which(NYC_transit_tidy$route_name == "A"))` distinct stations serve the A train.


Of the stations that serve the A train, how many are ADA compliant?

`r length(which((NYC_transit_tidy$route_name == "A") & (NYC_transit_tidy$ada == "TRUE")))` stations.